\documentclass{article}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{draftwatermark}

%\SetWatermarkText{DRAFT}
%\SetWatermarkScale{3}
%\SetWatermarkLightness{0.5}

\begin{document}

\title{EP 501 Homework 4:  Differentiation, Integration, and Multidimensional functions}

\maketitle

~\\
\textbf{Instructions:}  
\begin{itemize}
  \item Submit all source code and publish Matlab results in .pdf form via Canvas.  Please zip all contents of your solution into single file and then submit in a single zip file.    
  \item Discussing the assignment with others is fine, but you must not copy anyone's code.  
  \item I must be able to run your code and produce all results by executing a single top-level Matlab script, e.g. \textsf{assignment1.m} or similar.  
  \item You may use any of the example codes from our course repository:  \url{https://github.com/mattzett/EP501/}.
  \item Do not copy verbatim any other codes (i.e. any source codes other than from our course repository).  You may use other examples as a reference but you must write you own programs (except for those I give you).  

\end{itemize}
~\\~\\~\\
\textbf{Purpose of this assignment:}  
\begin{itemize}
  \item Deal with numerical differentiation to solve complex problems.  
  \item Develop good coding and documentation practices, such that your programs are easily understood by others.  
  \item Exercise good judgement in numerical problem setup.
  \item Demonstrate higher reasoning to synthesize a problem and devise a basic algorithm to solve it.  
\end{itemize}

\pagebreak

\begin{enumerate}
  \item  Vector derivatives and multidimensional plotting:  
  \begin{itemize}  
    \item[(a)] Plot the two components of the vector magnetic field defined by the piecewise function:
    \begin{equation}
      \mathbf{B}(x,y)= \left\{
      \begin{array}{c}
      \frac{\mu_0 I}{2 \pi a^2} \sqrt{x^2+y^2} \left( - \frac{y}{\sqrt{x^2+y^2}}\hat{\mathbf{e}}_x + \frac{x}{\sqrt{x^2+y^2}} \hat{\mathbf{e}}_y \right) \qquad \left( \sqrt{x^2+y^2} < a \right)
      \\
      \frac{\mu_0 I}{2 \pi \sqrt{x^2+y^2}} \left( - \frac{y}{\sqrt{x^2+y^2}} \hat{\mathbf{e}}_x + \frac{x}{\sqrt{x^2+y^2}} \hat{\mathbf{e}}_y \right) \qquad \left( \sqrt{x^2+y^2} \ge a \right)
      \end{array}
      \right.      
    \end{equation}
    Assume the parameters in this equation have the values:
    \begin{eqnarray}
      I &=& 10 \qquad \left( \mathrm{A} \right) \nonumber \\
      \mu_0 &=& 4 \pi \times 10^{-7} \qquad (\mathrm{H/m}) \nonumber \\
      a &=& 0.005 \qquad (\mathrm{m}) \nonumber 
    \end{eqnarray}
    Use an image plot (\texttt{imagesc}, \texttt{pcolor}) for each component and have your plot show the region $-3 a \le x \le 3 a, -3a \le y \le 3a$.  Make sure you add a colorbar and axis labels to your plot.  You will need to define a range and resolution in $x$ and $y$, and create a meshgrid from that.  Be sure to use a resolution fine enough to resolve important variations in this function.  
    \item[(b)]  Make a quiver plot of the magnetic field $\mathbf{B}$; add labels, etc.
    \item[(c)]  Compute the numerical curl of $\mathbf{B}$, i.e. $\nabla \times \mathbf{B}$.  Use centered differences on the interior grid points and first-order derivatives on the edges.  Plot your result using \texttt{imagesc}, or \texttt{pcolor}.  
    \item[(d)]  Compute $\nabla \times \mathbf{B}$ analytically.  Plot the  alongside your numerical approximation.  Demonstrate that they are suitably similar.  
    \item[(e)]  Compute and plot the scalar field:
%    \begin{equation}
%    \Phi(x,y,z)=\frac{q}{4 \pi \epsilon_0} \frac{e^{-\alpha \sqrt{x^2+y^2+z^2} }}{\sqrt{x^2+y^2+z^2}} \left(1+\frac{\alpha \sqrt{x^2+y^2+z^2}} {2} \right)
%    \end{equation}
    \begin{equation}
      \Phi(x,y,z)= \left\{
      \begin{array}{c}
      \frac{Q}{4 \pi \epsilon_0 a} - \frac{Q}{8 \pi \epsilon_0 a^3} \left(x^2+y^2+z^2 - a^2 \right)
      \qquad \left( \sqrt{x^2+y^2+z^2} < a \right)
      \\
      \frac{Q}{4 \pi \epsilon_0 \sqrt{x^2+y^2+z^2}} \qquad \left( \sqrt{x^2+y^2+z^2} \ge a \right)
      \end{array}
      \right.      
    \end{equation}
    Use the parameters:
    \begin{eqnarray}
      Q &=& 1  \qquad \left( \mathrm{C} \right) \nonumber \\
      a &=& 1 \qquad (\mathrm{m}) \nonumber \\
      \epsilon_0 &=& 8.854 \times 10^{-12} \qquad (\mathrm{F/m}) \nonumber 
    \end{eqnarray}    
    and plot this function in the region $-3a \le x \le 3a, -3a \le y \le 3a$ in the $z=0$ plane. Be sure to use a resolution fine enough to resolve variations in this function (aside from those associated with the singularity).
    \item[(e)]  Write a function to numerically compute the Laplacian of a scalar field, i.e. $\nabla^2 \Phi$.  Plot your result and include appropriate labels and colorbars.  
    
    
  \end{itemize}
  \item  Integration in multiple dimensions.  
  \begin{itemize}
    \item[(a)] Numerically compute the electrostatic energy in the region $R \equiv -3a \le x \le 3a, -3a \le y \le 3a, -3a \le z \le 3a$, defined by the integral:  
    \begin{equation}
      W_E = - \frac{1}{2} \iiint_R \left( \epsilon_0 \nabla^2 \Phi \right) \Phi ~dx dy dz
    \end{equation}
    using a iterated trapezoidal or multi-dimensional trapezoidal method.  
    \item[(b)]  Compute and plot the parametric path
    \begin{equation}
       \mathbf{r}(\phi) \equiv x(\phi)\hat{\mathbf{e}}_x + y(\phi)\hat{\mathbf{e}}_y = r_0 \cos \phi ~ \hat{\mathbf{e}}_x + r_0 \sin \phi ~ \hat{\mathbf{e}}_y \qquad (0 \le \phi \le 2 \pi)
    \end{equation}    
    in the $x,y$ plane on the same axis as your magnetic field components from the previous problem (plot the path in the figure for each component).  Take $r_0 = 2a$.  You will need to define a grid in $\phi$ to do this.
    \item[(c)]  Plot the two components of the magnetic field $\mathbf{B}(x(\phi),y(\phi))$ at the $x,y$ points along $\mathbf{r}$ (I recommend \texttt{yyaxis} for this) and visually compare against your image plots of the magnetic field.  
    \item[(d)]  Numerically compute the tangent vector to the path $\mathbf{r}$ by performing the derivative:  $d \mathbf{r} / d \phi$.  Compare your numerical results against the analytical derivative (e.g. plot the two) and refine your grid in $\phi$ (if necessary) such that you get visually acceptable results - i.e. such that the path appears circular.  
    \item[(e)]  Numerically compute the magnetic field integrated around the path $\mathbf{r}$, i.e.:
    \begin{equation}
       \frac{1}{\mu_0} \int_{\mathbf{r}} \mathbf{B} \cdot d \boldsymbol{\ell} 
    \end{equation}
    where the differential path length is given by:
    \begin{equation}
      d \boldsymbol{\ell} = \frac{d \mathbf{r}}{d \phi} d \phi 
    \end{equation}   
  \end{itemize}
\end{enumerate}

\end{document}
